---
title: "Problem Set 6, Stats 506 F19"
author: "Jie Cao"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(tidyverse)
library(kableExtra)
```

Here is the list of files I submitted for this problem set 6:

1. Question 1

  + ps6_q1.sas (SAS script)
  
  + ps6_q1.log (SAS log)
  
  + ps6_q1.csv (Summary table presenting results)
  
2. Question 2

  + ps6_q2.sas (SAS script)
  
  + ps6_q2.log (SAS log)
  
  + ps6_q2_b.csv (Summary table answering question 2b)
  
  + ps6_q2_c.csv (Summary table answering question 2c)
  
3. Question 3

  + ps6_q3.sas (SAS script)
  
  + ps6_q3.log (SAS log)
  
  + ps6_q3_b.csv (Summary table answering quetion 3b/2b)
  
  + ps6_q3_c.csv (Summary table answering question 3c/2c)
  
  
\pagebreak
  
### Question 1

Here's my solution text to question 1.

```{r q1_sas}
cat("```SAS\n")
writeLines(readLines("./ps6_q1.sas"))
cat("```\n")
```

```{r q1_table}
q1 = read_csv("./ps6_q1.csv")
cap1 = paste("Model summaries.", 
             "This table shows the relative effect (with 95\\% confidence intervals)", 
             "of the atypical condition on each of four curvature measures.", 
             "Standard deviations for each variance component -", 
             "subjects, exemplar, and error are also shown.")
q1 %>% 
  knitr::kable(cap = cap1, 
               format = "latex", 
               booktabs = TRUE,
               digits = 2) %>% 
  kableExtra::kable_styling(latex_options = "hold_position") %>% 
  kableExtra::row_spec(0, bold = TRUE)
```

\pagebreak

### Question 2

Here's my solution text to question 2. Answers to 2b and 2c are shown on the last page of this document. 

```{r q2_sas}
cat("```SAS\n")
writeLines(readLines("./ps6_q2.sas"))
cat("```\n")
```

\pagebreak

### Question 3

Here's my solution text to question 3. Answers to 3b and 3c (same as answers to 2b and 2c) are shown on the last page of this document. 

```{r q3_sas}
cat("```SAS\n")
writeLines(readLines("./ps6_q3.sas"))
cat("```\n")
```

\pagebreak

**(2b/3b) What is the national average home temperature at night, among homes that use space heating?**

```{r q3_b_table}
q3_b = read_csv("./ps6_q3_b.csv")
q3_b_pe = q3_b$`Average temperature at night`
q3_b_ci = q3_b$`95% Confidence Interval`
```

The national average home temperature (95% CI) at night, among homes that use space heating is `r q3_b_pe` `r q3_b_ci` $^\circ$F.

**(2c/3c) What is the average winter home temperatures at night, during the day with someone home, and during the day with no one home (when applicable), by census division?**

```{r q3_c_table}
q3_c = read_csv("./ps6_q3_c.csv")
cap3c = paste("National average temperature (95\\% CI)", 
              "during the day with someone home,", 
              "during the day with no one home,", 
              "and at night,", 
              "by census division")
q3_c %>% 
  knitr::kable(format = "latex", 
               booktabs = TRUE, 
               cap = cap3c) %>% 
  kableExtra::kable_styling(latex_options = "hold_position", 
                            full_width = FALSE) %>% 
  kableExtra::column_spec(2:4, width = "12em") %>% 
  kableExtra::row_spec(0, bold = TRUE)
```

