---
title: "Problem Set 3, Solutions"
author: "Stats 506, F19"
date: "November 12, 2019"
output: 
  html_document: 
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r run_scripts, include = FALSE}
source('./ps3_q1.R')
source('./ps3_q2.R')
```

## About

In the solutions below, we focus on the answers to part c of each question.
For functions please see the corresponding R script at the
course git [repo](https://github.com/jbhender/Stats506_F19/tree/master/problem_sets). 

## Question 1

a. See `confint_jackknife()` defined in `ps3_q1.R`.
a. See `confint_boot()` defined in `ps3_q1.R`.
a. The table below contains CIs for each method and dose.

```{r}
cap = "**Table 1.** *95% Confidence intervals for the ToothGrowth data.*"
tooth_growth_cis %>% 
  mutate( ci = sprintf('(%4.2f, %4.2f)', lower, upper) ) %>%
  pivot_wider( id_cols = "dose", values_from = 'ci', names_from = 'method') %>%
  knitr::kable( format = 'html', caption = cap ) %>%
  kableExtra::kable_styling("striped", full_width = TRUE)
```

## Question 2

a. See `confint_jackknife_mc()` defined in `ps3_q2.R`.
a. See `confint_boot_mc()` defined in `ps3_q2.R`.
a. In the Monte Carlo simulation shown here, we compare CIs constructed using 
the jackknife or one of three bootstrap methods for estimating the ratio of means
$E[X] / E[Y] = 2$ when 
$X_1, \dots, X_{30} \stackrel{_{iid}}{\sim} \textrm{Exponetial}(1)$ 
and $Y_1, \dots, Y_{20} \stackrel{_{iid}}{\sim} \textrm{Exponential}(0.5)$.

We use 10,000 Monte Carlo replicates and 10,000 bootstrap samples. The 
Monte Carlo replicates are split into 10 blocks of 1,000 to avoid allocating too
large of vectors in the bootstrap function.

The results are in the table below.

```{r pq3_q2c}
cap = paste0("**Table 2.** *Comparison of jaccknife and bootstrap confidence ",
 "intervals for a ratio of exponentials. *" )
result_tab %>%
  transmute( 
    Method = method,
    Coverage = sprintf('%4.2f (%4.2f, %4.2f)', cov_prob, 
                         cov_prob - qnorm(.975)*se_cov, 
                         cov_prob + qnorm(.975)*se_cov ),
    Length = sprintf('%4.2f (%4.2f, %4.2f)', length, 
                          length - qnorm(.975)*se_length, 
                          length + qnorm(.975)*se_length ),
    Shape = ifelse( method %in% c('basic bootstrap', 'jackknife'), '1 (-, -)', 
                      sprintf('%4.2f (%4.2f, %4.2f)', shape, 
                          shape - qnorm(.975)*se_shape, 
                          shape + qnorm(.975)*se_shape )
                      )
  ) %>%
  knitr::kable( format = 'html', caption = cap) %>%
  kableExtra::kable_styling("striped", full_width = TRUE)
```

