---
title: "Problem Set 2, Solutions"
author: "Stats 506, F19"
date: "10/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r run_scripts, include = FALSE}
source('./ps2_q1.R')
source('./ps2_q2.R')
```

## About

In the solutions below, we focus on the answers to specific content questions.
For code and data manipulations please see the corresponding R script at the
course git [repo](https://github.com/jbhender/Stats506_F19/). 

## Question 1

In this question, we use the 2015 RECS data to answer questions about home
heating behavior. 

### (a) What is the national average home temperature at night in winter, among homes that use space heating?

The national average home temperature at night, in winter, among homes that use space heating is `r str_temp_night` °F.

### (b) Space heating fuel used, by division and urban type. 

```{r q1b_table, fig.cap = cap}
fuel_types_wide = rename(fuel_types_wide, 
                         Division = division, 
                         `Urban Type` = urban)
DT::datatable( fuel_types_wide  )
cap = paste0(
"**Table 1.** *Percents of home using different types of ",
"space heating fuels.* Values are shown as percents with 95% confidence",
"intervals in parentheses."
)
```

### (c) Comparing day and night temperatures, by division and urban type.

```{r q1c_figure, fig.cap = cap}
temps %>%
  ungroup() %>%
  mutate( division = factor(division, levels = div_order),
          `Temperature Type` = type
  ) %>%
  ggplot( aes(y = est, x = division, color = `Temperature Type`, 
              group = type) ) +
  geom_point( position = position_dodge(width = .5), alpha = 0.5 ) +
  geom_errorbar( aes(ymin = lwr, ymax = upr), alpha = 0.5, 
                 position = position_dodge(width = .5)
  ) + 
  coord_flip() +
  facet_wrap(~urban) + 
  theme_bw() +
  scale_color_manual( values = c('goldenrod', 'darkorange', 'black')) +
  ylab( expression('Temperature ' (degree*'F')) ) +
  ylim( c(60, 75)) +
  xlab('')

cap = paste0(
"**Figure 1.** *Average household temperatures in winter.* This plot ",
"compare winter home temperatures of three types by divison and urban ", 
"type. Generally, temperatures are warmest during the day when someone is ",
"home, cooler at night, and coolest during the day with no one at home."
)
```

### (d) National, median differences between day and night temperatures by heating equipment behavior 

```{r q1d_plot, fig.cap = cap}
temp_se %>%
  mutate( heat_type = factor( heat_behavior, hb_order, hb_labels) ) %>%
  ggplot( aes( y = heat_type, x = est) ) +
  geom_point() +
  geom_errorbarh( aes(xmin = lwr, xmax = upr)) +
  theme_bw() +
  ylab("Heating Equipment Behavior") + 
  xlab( 
    expression("Median difference between day and night temperatures " (degree*'F'))
  )
cap = paste0(
 "**Figure 2.** *Median national differences in day and night temperatures.* ",
 "This figure shows the median national difference between household ",
 "temperatures in winter during the day, with somoone home, and at night. The ",
 "median differnce is largest, at 3 ºF, among those using a programmable ",
 "thermostat.  The estimated standard errors are zero for the most commont ",
 "behvarior types as the distribution of temperature differences has a ",
 "large point mass near the median for these types."
)
```

***

## Problem 2
In this question, we analyze (computer) mouse tracking experiments to show that 
"atypical" examples of various animal species are associated with more 
curvature on the path to the correct reponse. These curvature measures are a
proxy for indecision or cognitive burden.  

We analyze these data using linear mixed models with condition (typical vs
atypical) as the sole covariate and random intercepts for subject and examplar
to account for the repeated measurs nature of the data.  Subject level (random)
intercepts help to account for subject-to-subject differences in curvature 
irrespective of condition.  Similarly, exmemplar
(random) intercepts help to account for the fact that exemplars are repeated 
across subjects.

Based on the results below, in this experiment the atypcial condition had
the largest effect on the average absolute deviation measure which was nearly
twice as large, on average, in the atypical condition.  However, comparing
confidence intervals for the relative effects of condition on each curvature
measure, the only statistically meaningful difference is between average absolute
deviation and total distance.  

```{r q2_plot, fig.cap = cap}
ci_df %>%
  arrange( est ) %>%
  mutate( measure = factor(measure, levels = unique(measure) ) ) %>%
  ggplot( aes( y = measure, x = est) ) +
  geom_point() +
  geom_errorbarh( aes(xmin = lwr, xmax = upr) ) + 
  geom_vline( xintercept = 1, lty = 'dashed') + 
  xlab('relative effect on curvature measure of an atypical exemplar') + 
  ylab('') +
  theme_bw()
cap = paste(
  "**Figure 3.** *Relative effect of atypical condition on each curvature measure.*"
)
```

```{r q2_table}
cap = paste(
  "**Table 2.** *Model summaries.* This table shows the relative effect (with ",
  "95% Wald confidence intervals) of the atypical condition on each of four ",    
  "curvature measures. ",
  "Standard deviations for each variance component - subjects, exemplar, and ",
  "error - are also shown with 95% confidence intervals based on the profile ", 
  "likelihood."
)

ci_df %>% 
  select(measure:Error) %>% 
  knitr::kable( caption = cap )
```

