---
title: "Problem Set 1, Solutions"
author: "Stats 506, F19"
date: "`r format( Sys.Date(), '%B %d, %Y')`"
output: pdf_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Example solutions for the first problem set are below. The
source file for this document and underlying scripts can all be found 
at the course git repo: https://github.com/jbhender/Stats506_F19/.

## Problem 1
Here is the solution text for `ps1_q1.sh`. See the `stats506_F19` repo for an executable version.

```{r ps1_q1, results = 'asis'}
cat('```bash\n')
writeLines(readLines('./ps1_q1.sh'))
cat('```')
```

\pagebreak

## Problem 2
The text of the script is below. See the `stats506_F19` repo for an executable 
version `cutnames.sh`

```{r cutnames, results = 'asis'}
cat('```bash\n')
writeLines(readLines('./cutnames.sh'))
cat('```')
```

\pagebreak

## Problem 3

### Parts a, b, c, and d
Below is the code form `ps1_q3.R` answering parts a-d.
```{r q3_rfuncs, results = 'asis'}
cat('```R\n')
writeLines( readLines('./ps1_q3.R'))
cat('```\n')
```

### Part e

Here are the measures computed on the test data,
presented as a nicely formatted table.

```{r test_measures, echo = 'TRUE', message = FALSE}
source('./ps1_q3.R')
cap = 
 '*Measures of curvature computed from the "test" trajectories.*'
knitr::kable(test_meas, digits = 1, col.names = cn, caption = cap)
```


